---
title: "Untitled"
author: "Daniela Kemp"
date: "2023-12-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Installation and Package loading

```{r Packages, message=FALSE, warning=FALSE, include=FALSE}
library("ggplot2")
library('dplyr')
library('Rlab')
library('stringr')
```
ALL THE FOLLOWING FUNCTIONS, UP TO AND FUSION SHOULD ALSO HAVE A TABLE EITHER AS INPUT, OR WORK ON A TABLE THAT IS CHANGED THROUGH THEIR CALL.

#Population Change Function to model population development in one year, given parameters:

matrix, parameters-

```{r Population Change}
population_change <- function(matrix,parameters, ...) {
    N<-tail(matrix[,"overall"],1)
    Fec     <- parameters$l_max / (1 + parameters$b *N )      
    indFec  <- rpois(n=min(N,200), lambda=Fec) 
    #indFec  <- rpois(n=N, lambda=Fec) 

    new_Generation<-N*mean(indFec)
  return(new_Generation)
}

```

The population_dependent option in dispersal has different values depending on the ratio of current population/stable population. Here, we use a transformed ratio of the ratio current population/max population and take the ratio of random realisations above that as a  percentage. This results in dispersal rates close to zero for small populations, but more varied and larger rates for populations close, or above the maximum stable population size. Due to a linear factor, the results all fall between 0 and 0.4.

```{r Testing population dependent dispersal}
to_try<-seq(50,650,25)/500
empty<-matrix(data=NA,nrow=20,ncol=length(to_try))
colnames(empty)<-to_try
for(i in 1:dim(empty)[1]){
  k<-0
  for (z in to_try){
    k<-k+1
    test_scale<-((1.1*z)**3)*1.7
    empty[i,k]<-(sum(rlogis(50, location=2.5)<test_scale)/50)*0.25
  }
}
boxplot(empty, main="Dispersal probability vs \n current/sustainable population")
```
#Dispersal Function to model dispersal in one patch in one year given parameters:

disp_prob - probability to change the patch

Output: two matrices of one row, both showing the new population

```{r Dispersal}
randomly_drawn<-function(n,p){
  return(sum(rbern(n,p)))
}

return_prob<-function(n,N_max,max_disp){
      limit<-((1.1*n/N_max)**3)*1.7
      return(sum(rlogis(50, location=2.5)<limit)/50)*max_disp
}

dispersal<-function(matrix1, matrix2, parameter1,parameter2, year, method="default",  ...){
  last_row1<-tail(matrix1,n=1)
  last_row2<-tail(matrix2,n=1)
  n1<-last_row1[,"overall"]
  n2<-last_row2[,"overall"]

  pop_max1<-  (parameter1$l_max-1)/parameter1$b
  pop_max2<-  (parameter2$l_max-1)/parameter2$b
  
  disp_prob1<-case_when(method=="even" ~ parameter1$dispersal_prob,
                        method=="population_dependent" ~ return_prob(n1,pop_max1,parameter1$dispersal_prob)*parameter1$wanderlust,
                        TRUE ~ 0.1)
  
  disp_prob2<-case_when(method=="even" ~ parameter2$dispersal_prob,
                        method=="population_dependent" ~ return_prob(n2,pop_max2,parameter2$dispersal_prob)*parameter2$wanderlust,
                        TRUE ~ 0.1)
  
  d_1to2<-floor(last_row1[,c("AA","AB","BB")]*disp_prob1)
  d_2to1<-floor(last_row2[,c("AA","AB","BB")]*disp_prob2)
  
  change_1<-c(d_2to1-d_1to2,sum(d_2to1)-sum(d_1to2))
  change_2<-c(d_1to2-d_2to1,sum(d_1to2)-sum(d_2to1))
  
  new_row1<-matrix(data=c(last_row1+change_1),ncol = 4, dimnames = list(paste("Year_",year,sep=""),colnames(matrix1)))
  
  new_row2<-matrix(data=c(last_row2+change_2),ncol = 4, dimnames = list(paste("Year_",year,sep=""),colnames(matrix1)))
  return(list(row1=new_row1,row2=new_row2))
}
```
#Gamete Production Individual Model the gamete production of one individual. Parameters:

Patch - to see how well-fitted the individual is Genotype - [A,B,M]\<-AA,BB,Mixed:AB, to see what kind of gametes the individual can produce, and to check for fitness. f_max - how many gametes the individual would produce under optimal circumstances. h - dominance coefficient, default set to 0.5 s - selection coefficient

#Gamete Production population: models the gamete production of one entire population, produces list of gametes.Parameters:

Patch - which patch, has to be given to gamete individual Genotypes - a ratio, than in combination with number of indiviuals, or a count for all genotypes

Output: an aggregated list/vector of all Gametes

```{r Gamete Population}
gamete_production<-function(matrix, parameters,N_new, method="default", ...){

unfavored<-ifelse(parameters$favored=="AA", "BB","AA")
favored<-parameters$favored

prob_unfavored<-(1-parameters$s)
prob_heterozygen<-(1-parameters$h*parameters$s)
prob_favored<-1

fmax<-20
favored_mult<-ceiling(2*matrix[,favored]*prob_favored*fmax)
unfavored_mult<-ceiling(2*matrix[,unfavored]*prob_unfavored*fmax)
heterozygen_mult<-ceiling(matrix[,"AB"]*prob_heterozygen*fmax)

gamete_vector<-rep(substr(favored,1,1),favored_mult)
gamete_vector<-c(gamete_vector,rep(substr(unfavored,1,1),unfavored_mult))
gamete_vector<-c(gamete_vector,rep("A",heterozygen_mult),rep("B",heterozygen_mult))
return(gamete_vector)
}
```

#Fusion models the fusion of gametes in one patch. Parameters:

Gamete_List - a vector/list of gametes, from which two gametes are randomly sampled until N individuals are created reached. N - number of individuals that should be in the patch at the end

Output: 3 numbers, describing the number of individuals in a patch with the genotypes A,B,M

```{r Fusion}
test_vector<-c("A","A","A","A","A","B","B","B","B","B","A","A","A","A","A","B","B","B","B","B","A","A","A","A","A","B","B","B","B","B")
test_N<-12

Fusion<-function(gamete_vector=test_vector,N=test_N){ 
  A<-0
  B<-0
  M<-0
  num_gametes<-length(gamete_vector)
  
  #the vector of gametes is randomly sampled in-place
  ifelse(N*2>num_gametes,
    final_gametes<-gamete_vector[sample(num_gametes)],     
    final_gametes<-gamete_vector[sample(num_gametes,size=N*2,replace=FALSE)])

    for (i in seq(1,length(final_gametes)-1,2)){
    first<-final_gametes[i]
    second<-final_gametes[i+1]
    ifelse(first==second,
           ifelse(first=="A",
                  A<-A+1,
                  B<-B+1),
           M<-M+1)
  }
  return(c(A,M,B,A+B+M))
}
```
Set-up initialize two lists, both containing a matrix, and a list of parameters N0 should be dividable by 4. #Combined models, given a number of years, the number of individuals of different genotypes in each patch. Parameters:

```{r}
#Set values for selection_coefficient, heterozygote_coefficient, if they are symmetrical
selection_coefficient <- 0#.01
heterozygote_coefficient <-0#.5
disp_coefficient<-0.05

para_patch1<-list(N0=400, b=0.001, l_max=1.4, 
                  wanderlust=1, dispersal_prob=disp_coefficient,
                  favored="AA", s=selection_coefficient,h=heterozygote_coefficient)

para_patch2<-list(N0=400, b=0.001, l_max=1.4,
                  wanderlust=1, dispersal_prob=disp_coefficient,
                  favored="BB", s=selection_coefficient,h=heterozygote_coefficient)

matrix_patch1<-matrix(data=c(para_patch1$N0/4,para_patch1$N0/2,para_patch1$N0/4,para_patch1$N0),ncol=4,dimnames = list(c("Year_0"),c("AA","AB","BB","overall")))

matrix_patch2<-matrix(data=c(para_patch2$N0/4,para_patch2$N0/2,para_patch2$N0/4,para_patch2$N0),ncol=4,dimnames = list(c("Year_0"),c("AA","AB","BB","overall")))

Patch_1_list<-list(parameters=para_patch1, matrix=matrix_patch1)
Patch_2_list<-list(parameters=para_patch2, matrix=matrix_patch2)
```


```{R Definition of combined function}

run_several<-function(repetitions,years,paras1,paras2,one_result=FALSE){
  patch1_result<-matrix(nrow=years+1,ncol=0)
  patch2_result<-matrix(nrow=years+1,ncol=0)
  
  for (i in 1:repetitions){
  matrix_patch1<-matrix(data=c(paras1$N0/4,paras1$N0/2,paras1$N0/4,paras1$N0),ncol=4,dimnames = list(c("Year_0"),c("AA","AB","BB","overall")))
  matrix_patch2<-matrix(data=c(paras2$N0/4,paras2$N0/2,paras2$N0/4,paras2$N0),ncol=4,dimnames = list(c("Year_0"),c("AA","AB","BB","overall")))
  
  Patch_1_list<-list(parameters=paras1, matrix=matrix_patch1)
  Patch_2_list<-list(parameters=paras2, matrix=matrix_patch2)

  for(i in 1:years){
    N1_new<-population_change(Patch_1_list$matrix,Patch_1_list$parameters)
    N2_new<-population_change(Patch_2_list$matrix,Patch_2_list$parameters)
    
    disp_result<-dispersal(matrix1=Patch_1_list$matrix,matrix2=Patch_2_list$matrix, parameter1=Patch_1_list$parameters, parameter2=Patch_2_list$parameters, year=i,method="even")
    
    gamete_vector1<-gamete_production(disp_result$row1,Patch_1_list$parameters, N1_new)
    gamete_vector2<-gamete_production(disp_result$row2,Patch_2_list$parameters, N2_new)
    
    result1<-matrix(data=Fusion(gamete_vector = gamete_vector1, N=N1_new),ncol=4,dimnames = list(paste("Year_",i,sep=""),c("AA","AB","BB","overall")))
    
    result2<-matrix(data=Fusion(gamete_vector = gamete_vector2, N=N2_new),ncol=4,dimnames = list(paste("Year_",i,sep=""),c("AA","AB","BB","overall")))
    
    Patch_1_list$matrix<-rbind(Patch_1_list$matrix,result1)
    Patch_2_list$matrix<-rbind(Patch_2_list$matrix,result2)
    
  }
  if (one_result){
  patch1_result<-cbind(patch1_result,Patch_1_list$matrix,Patch_2_list$matrix)}
  else{
    patch1_result<-patch1_result<-cbind(patch1_result,Patch_1_list$matrix)
    patch2_result<-patch2_result<-cbind(patch2_result,Patch_2_list$matrix)
  }
  }
  ifelse(one_result,
         return(patch1_result),
         return(list(patch1_result,patch2_result)))
  
}

```

```{r}
set.seed(11)
#Minimal example, two Patches, no advantages for either genotype, 200 years, 400 individuals is the stable max population, no dispersal, how does the amount of individuals with different genotypes change? 

selection_coefficient <- 0
heterozygote_coefficient <-0

para_patch1<-list(N0=400, b=0.001, l_max=1.4, 
                  wanderlust=1, dispersal_prob=0,
                  favored="AA", s=selection_coefficient,h=heterozygote_coefficient)

para_patch2<-list(N0=400, b=0.001, l_max=1.4,
                  wanderlust=1, dispersal_prob=0,
                  favored="BB", s=selection_coefficient,h=heterozygote_coefficient)
Big_Patches<-run_several(repetitions = 10, years=200, paras1 = para_patch1, paras2=para_patch2, one_result = TRUE)

```


```{r Plot Def}
many_plotted_lines<-function(to_plot,upper_y=500,to_vis=list(AA=TRUE,AB=TRUE,BB=TRUE,overall=TRUE,legend=FALSE),title="", axes=TRUE){
  
  column_names<-c("AA","AB","BB","overall")
  plot(to_plot[,"overall"], col=alpha("black",0), lty=1, lwd=1, las=1, type="l", pch=16, xlab="generations", ylab="individuals (N)", ylim=c(0,upper_y),main=title,axes=axes) # 
  colors<-c("red","violet","blue","black")
  for (i in 1:dim(to_plot)[2]){
    which_type<-str_detect(colnames(to_plot)[i],column_names)
    if(sum(which_type)>0){
      yes_no<-to_vis[[column_names[which_type]]]
      if (yes_no){
        lines(to_plot[,i], lty=1, col=alpha(colors[which_type],0.2), lwd=1)
        }
    }
  }
  
  if(to_vis$AA){
    mean_AA_1<-apply(to_plot[,str_detect(colnames(to_plot),"AA")],1, mean)
    lines(mean_AA_1,col="darkred",lwd=2,lty=3)}
  if(to_vis$AB){
    mean_AB_1<-apply(to_plot[,str_detect(colnames(to_plot),"AB")],1, mean)
    lines(mean_AB_1,col="darkviolet",lwd=2,lty=3)}
  if(to_vis$BB){
    mean_BB_1<-apply(to_plot[,str_detect(colnames(to_plot),"BB")],1, mean)
    lines(mean_BB_1,col="darkblue",lwd=2,lty=3)}
  if(to_vis$overall){
    mean_overall<-apply(to_plot[,str_detect(colnames(to_plot),"overall")],1, mean)
    lines(mean_overall,col="black",lwd=2,lty=3)}
  
  if(to_vis$legend){
  legend("topright",
    c("genotype AA", "genotype AB", "genotype BB"),inset=c(0,0.00), lty=1, lwd=2, cex=0.5, col=c(alpha("red",0.4), alpha("violet",0.4),alpha("blue",0.4)), box.col=NA, horiz = TRUE, xpd=TRUE)
  legend("topright",
    c("mean AA", "mean AB", "mean BB", "mean overall"),inset=c(0,-0.05), lty=3, lwd=2, cex=0.5, col=c("darkred", "darkviolet","darkblue","black"), box.col=NA, horiz = TRUE, xpd=TRUE)}
}
```

```{r Visualization Baseline}
#In General, in this population size, genotypes are not safe from extinction due to random chance- see columns 307 to 40, in the resulting patches, where  BB frequently drops to 0, and AB is mostly in the low double digits later on.
#to_plot<-Big_Patches
many_plotted_lines(Big_Patches,to_vis=list(AA=TRUE,AB=TRUE,BB=TRUE,overall=TRUE,legend=TRUE),title="Genotype Frequency \n no selection, no dispersal",upper_y = 550)
tail(Big_Patches[,37:40],10)

```

```{r}
set.seed(25)
start_time <- Sys.time()
#Low selection coefficient, low chance of dispersal, 300years, max population of 500, same circumstances in both patches. To improve time-performance, gamete production does not use random draws, but prob*Gametes

selection_coefficient <- 0.05
heterozygote_coefficient <-0.5

para_patch1<-list(N0=200, b=0.001, l_max=1.5, 
                  wanderlust=1, dispersal_prob=0.05,
                  favored="AA", s=selection_coefficient,h=heterozygote_coefficient)

para_patch2<-list(N0=200, b=0.001, l_max=1.5,
                  wanderlust=1, dispersal_prob=0.05,
                  favored="BB", s=selection_coefficient,h=heterozygote_coefficient)

two_tables<-run_several(repetitions = 15, years=500, paras1 = para_patch1, paras2=para_patch2)

Small_changes_1<-two_tables[[1]]
Small_changes_2<-two_tables[[2]]
run_time <- Sys.time()-start_time

```

```{r Visualization small Advantages, small Dispersal}
#Population and time were somewhat increased, looking at the number of AA(favored), and AB(97.5% as favored) genotypes, to see whether this shows a significant change over 10 experiments and 300years
many_plotted_lines(Small_changes_1,upper_y=450, to_vis=list(AA=TRUE,AB=TRUE,BB=FALSE,overall=FALSE,legend=FALSE),title="AA-favored patch \n individuals with an A gene")
legend("topright",
  c("AB individuals", "AA individuals", "mean AB", "mean AA"), lty=c(1,1,3,3), lwd=c(3,3,2,2), cex=0.8, col=c(alpha("violet",0.4),alpha("red",0.4), "darkviolet","darkred"), box.col=NA, horiz = TRUE, xpd=TRUE)
```

```{r Visualization small Advantages, small Dispersal}
#Population and time were somewhat increased, looking at the number of BB(favored), and AB(97.5% as favored) genotypes, to see whether this shows a significant change over 10 experiments and 300years
many_plotted_lines(Small_changes_2,upper_y=550, to_vis=list(AA=FALSE,AB=TRUE,BB=TRUE,overall=FALSE,legend=FALSE),title="BB-favored patch \n individuals with a B gene")

legend("topright",
  c("AB individuals", "BB individuals", "mean AB", "mean BB"), lty=c(1,1,3,3), lwd=c(3,3,2,2), cex=0.8, col=c(alpha("violet",0.4),alpha("blue",0.4), "darkviolet","darkblue"), box.col=NA, horiz = TRUE, xpd=TRUE)

```

Looking at these values, it seems that with a small enough selection coefficient, like 5% here, random chance will be in a lot of cases more influential than adaption, as we barely see any trend in one direction or the other. A random (series of) event(s) that may influence the frequency of genotypes seems to weight stronger than the adaptation.

Looking at these values, it seems that with a small enough selection coefficient, like 5% here, random chance will be in a lot of cases more influential than adaption, as we barely see any trend in one direction or the other. A random (series of) event(s) that may influence the frequency of genotypes seems to weight stronger than the adaptation.

```{r Parameter space exploration 1h+!, eval=FALSE, include=FALSE}
set.seed(33)
#Low selection coefficient, low chance of dispersal, 300years, max population of 500, same circumstances in both patches. To improve time-performance, gamete production does not use random draws, but prob*Gametes
list_of_matrices_1<-list()
list_of_matrices_2<-list()

heterozygote_coefficient <-0.5
selection_coefficients<-seq(0,6)/40
disperal_probs<-c(seq(0,5)/40,0.5)

for(sc in selection_coefficients){
  for (d in disperal_probs){
    para_patch1<-list(N0=200, b=0.001, l_max=1.5, 
                      wanderlust=1, dispersal_prob=d,
                      favored="AA", s=sc,h=heterozygote_coefficient)
    
    para_patch2<-list(N0=200, b=0.001, l_max=1.5,
                      wanderlust=1, dispersal_prob=d,
                      favored="BB", s=sc,h=heterozygote_coefficient)
    
    two_tables<-run_several(repetitions = 2, years=10, paras1 = para_patch1, paras2=para_patch2)

    Name1<-paste("P1_d=",d,"_sc=",sc,sep="")
    Name2<-paste("P2_d=",d,"_sc=",sc,sep="")
    list_of_matrices_11[[Name1]]<-two_tables[[1]]
    list_of_matrices_22[[Name2]]<-two_tables[[2]]
  }
}  
```

```{r Writing/Loading of computations}
#saveRDS(list_of_matrices_1, file="list_of_matrices_1.RData")
#saveRDS(list_of_matrices_2, file="list_of_matrices_2.RData")

list_of_matrices_1<-readRDS("list_of_matrices_1.RData")
list_of_matrices_2<-readRDS("list_of_matrices_2.RData")
```

```{r Visualization Parameterspace AA, fig.height=7, fig.width=14}
#Plotting an overview of the whole parameter space, for Patch 1.
par(cex=0.7, mai=c(0.06,0.06,0.06,0.06), oma=c(5,1,4,0))
#layout(matrix(c(1:45), nrow = 5, byrow = FALSE))
layout(matrix(c(1:49), nrow = 7, byrow = FALSE))

colors<-c("red","violet","blue","black")

for (name in names(list_of_matrices_11)){
  many_plotted_lines(to_plot=list_of_matrices_11[[name]],axes=FALSE, upper_y = 550)
  axis(1, at=c(0,100,200,300,400,500), labels = FALSE,tck = 0.02)
  axis(2,at=c(0,100,200,300,400,500), labels=FALSE,tck = 0.02)
}
mtext("AA-favored Patch, frequency of genotypes",side=3,cex=1.2, line= 2,outer=TRUE) 


mtext("selection coefficient:0.00                                  0.025                                                 0.05                                                     0.075                                    0.1                                                   0.125                                                0.15                ",                   
      side = 3,
      cex=0.8,
      outer = TRUE)
mtext("         dispersal prob.: 0.5               0.0125             0.1                0.075                0.05                 0.025                   0.0                    ",
      side=2, 
      cex=0.8,
      outer=TRUE )
par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
   plot(0, 0, type = 'l', bty = 'n', xaxt = 'n', yaxt = 'n')
legend("bottom",
  c("genotype AA", "genotype AB", "genotype BB"),inset=c(0,0.04), lty=1, lwd=4, cex=1.1, col=c(alpha("red",0.4), alpha("violet",0.4),alpha("blue",0.4)), box.col=NA, horiz = TRUE, xpd=TRUE)
legend("bottom",
  c("mean AA", "mean AB", "mean BB", "mean overall"),inset=c(0,-0.00), lty=3, lwd=4, cex=1.1, col=c("darkred", "darkviolet","darkblue","black"), box.col=NA, horiz = TRUE, xpd=TRUE)
```

```{r Visualization Parameterspace AA, fig.height=7, fig.width=14}
#Plotting an overview of the whole parameter space, for Patch 1.
par(cex=0.7, mai=c(0.06,0.06,0.06,0.06), oma=c(5,1,4,0))
#layout(matrix(c(1:45), nrow = 5, byrow = FALSE))
layout(matrix(c(1:49), nrow = 7, byrow = FALSE))

colors<-c("red","violet","blue","black")

for (name in names(list_of_matrices_2)){
  many_plotted_lines(to_plot=list_of_matrices_2[[name]],axes=FALSE, upper_y = 550)
  axis(1, at=c(0,100,200,300,400,500), labels = FALSE,tck = 0.02)
  axis(2,at=c(0,100,200,300,400,500), labels=FALSE,tck = 0.02)
}
mtext("BB-favored Patch, frequency of genotypes",side=3,cex=1.2, line= 2,outer=TRUE) 


mtext("selection coefficient:0.00                                  0.025                                                 0.05                                                     0.075                                    0.1                                                   0.125                                                0.15                ",                   
      side = 3,
      cex=0.8,
      outer = TRUE)
mtext("         dispersal prob.: 0.5               0.0125             0.1                0.075                0.05                 0.025                   0.0                    ",
      side=2, 
      cex=0.8,
      outer=TRUE )
par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
   plot(0, 0, type = 'l', bty = 'n', xaxt = 'n', yaxt = 'n')
legend("bottom",
  c("genotype AA", "genotype AB", "genotype BB"),inset=c(0,0.04), lty=1, lwd=4, cex=1.1, col=c(alpha("red",0.4), alpha("violet",0.4),alpha("blue",0.4)), box.col=NA, horiz = TRUE, xpd=TRUE)
legend("bottom",
  c("mean AA", "mean AB", "mean BB", "mean overall"),inset=c(0,-0.00), lty=3, lwd=4, cex=1.1, col=c("darkred", "darkviolet","darkblue","black"), box.col=NA, horiz = TRUE, xpd=TRUE)
```
