---
title: "Project"
output: html_notebook

---

Installation and Package loading

```{r}
library("ggplot2")
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or
by pressing *Ctrl+Alt+I*.

#Population Change Function to model population development in one year,
given parameters:

N0 - Number of adults in the patch 
b - Growth rate 
l_max - maximum population that may survive in the habitat 
patch - on which patch we are working

OR: we only take Patch as input, we have pre-defined the parameters for
each patch, and do a short look up, and use the last row of the patch X
column to than compute the new population

```{r Population Change}

  Population <- function(N0=10, b=0.001, l_max=1.4, years=50) {
  ADLT           <- matrix(1, ncol=1, nrow=N0)     
  
colnames(ADLT) <- "patch"                        
  sumStats      <- NULL                            
  sumStats      <- rbind(sumStats, c(0,N0))        
  
  for(t in 1:years){
    
    N       <- nrow(ADLT)                           
    Fec     <- l_max / (1 + b * N)                      
    indFec  <- rpois(n=N, lambda=Fec)               
    OFF     <- matrix(1, ncol=1, nrow= sum(indFec)) 
    colnames(OFF) <- "patch"                        
    ADLT     <- OFF  
    
  if (t >= quota_T) {
      harvesting <- min(quota_N, nrow(ADLT))
      ADLT <- ADLT[sample(1:nrow(ADLT), size = nrow(ADLT)-harvesting, replace = F),] 
      
      }
    sumStats <- rbind(sumStats, c(t,nrow(ADLT))) 
}
    colnames(sumStats) <- c("year", "size")
  return(sumStats)

}

```

#Dispersal Function to model dispersal in one patch in one year given
parameters:

Patch - which patch are we looking at disp_prob - probability to change
the patch

```{r Dispersal}

```

#Gamete Production Function to model gamete production for all
individuals in one patch.

```{r}

```
